Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2020-01-07 paper/manuscript.tex
fixing figure captions and citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}
\maketitle

\begin{abstract}
Alchemical free energy calculations are a useful tool for predicting free energy differences associated with the transfer of small molecules from one environment to another.
The hallmark of these methods is the use of "bridging" potential energy functions representing \emph{alchemical} intermediate states that do not exist in as real chemical species. The data collected from these bridging alchemical thermodynamic states allows the computation of transfer free energies (or differences in transfer free energies) with orders of magnitude less simulation time than simulating the transfer process directly.

While these methods are highly flexible, care must be taken in avoiding common pitfalls to ensure that computed free energy differences can be robust and reproducible for the chosen force field, and that appropriate corrections are included to permit direct comparison with experimental data.

In this paper, we review current best practices for several popular application domains of alchemical free energy calculations, including relative and absolute small molecule binding free energy calculations to biomolecular targets.
\todo[inline, color=green!20]{JDC: Should we migrate this document to \url{https://github.com/alchemistry}? ASJSM: Let's migrate with the submitted version, but leave it here for now?}
\end{abstract}

 \end{frontmatter}

@@

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig1_what_is_alchemy/Figure.pdf}   
    \caption{\textbf{Illustration of the types of free energies differences that can be calculated using alchemical free energy methods.} A: Change in free energy due to a conformational change of the molecule. B: Partition coefficient such as $\log P$ or $\log D$ can be computed using alchemical free energy methods. Here, as an example the partition coefficient between methanol and water is shown. C: Free energy difference associated with the insertion of a molecule into a membrane. D: Effect of mutations of protein or host residues on free energies of binding.  E: Absolute free energies of binding of a small molecule to a host (e.g. protein), F: Relative free energy of binding of one molecule with respect to another, here toluene and benzyl alcohol.
    \label{fig:fig_what_is_alchemy}
    }
\end{figure}

%%%%%%%%%%%%%%%%%%%%
% Prerequesites    %
%%%%%%%%%%%%%%%%%%%%
\section{Prerequisites and Scope}
\label{sec:pre}
%Here you would identify prerequisites/background knowledge that are assumed by your work and your checklist which you view as critical, ideally giving links to good sources on these topics.
%Checklists are normally focused on errors made by users with training and experience in molecular simulations, so you can assume a basic familiarity with the fundamentals of molecular simulations.
\todo[inline, color=green!20]{ASJSM: @everyone This section is up for review to also check flow with intro and rest of document.}
This guide is meant to serve as a comprehensive starting point for new practitioners for free energy calculations.  It assumes moderate experience with molecular simulation packages and concepts, however.  It is also intended to serve as a reference guide with a convenient checklist (Section~\ref{sec:checklist}) for more seasoned practitioners, to help ensure their simulation meet standard alchemical simulation and analysis practices. It can also serve as a set of best practices to ensure simulation robustness and reproducibility which reviewers may wish to consider as they evaluate papers.

This guide aims to answer the following questions:
\begin{itemize}
    \item Is my problem suitable for an alchemical free energy calculation? 
    \item How do I choose the right free energy protocol and run it? 
    \item How do I analyze my simulations accurately? 
    \item What software tools are available to perform alchemical free energy calculations? 
\end{itemize}

We assume the reader possesses a basic familiarity with the principles of molecular mechanics, molecular dynamics simulations, statistical mechanics, and the biophysics of protein-ligand association.
For readers unfamiliar with these concepts that have never set up and run an equilibrium molecular dynamics simulation, we would recommend the following texts for getting started with these topics first. 
\begin{itemize}
    \item Statistical mechanics and thermodynamics background~\cite{hill1956statistical, alavi2011statistical}
    \item Understanding theory of how molecular simulations work~\cite{frenkel2002understanding}
    \item Docking reviews~\cite{yuriev2013latest, meng2011molecular} and software overview~\cite{pagadala2017software} and generating valid molecular structures ~\cite{loeffler2015fesetup}
    \item Getting started with setting up and running equilibrium molecular dynamics simulations: Gromacs tutorial~\cite{lemkul2018From}, Amber tutorial~\cite{amber}, HTMD: high-throughput Molecular Dynamics for Molecular Discovery~\cite{doerr2016htmd}*, or with VMD and NAMD~\cite{ribeiro2016easy}*
\end{itemize}
References marked with an asterisk (*) indicate nonessential reading or tutorials that provide further depth and understanding.

As some of the theoretical background can seem daunting, we provide an essentials guide to the theory behind alchemical free energy calculations in Section~\ref{sec:theory}.
In the remainder of this paper, we will cover topics that are key to the preparation~(\ref{sec:step1}), choice and use of correct protocols~(\ref{sec:step2}), and finally the best practices that should be used for the analysis of AFE calculations~(\ref{sec:step3}). 
Particular focus will be given to aspects of the molecular simulations which are unique to alchemical calculations---these include the calculation of transfer free energies (hydration free energies, partition coefficients, etc.), and binding free energies (absolute and relative).

While we try to address as many methods and practices as possible, the field of free energy calculations is broad, and there are many advanced topics that are left to future best practices documents focusing on specific issues. 
Below, we provide a non-exhaustive list of topics we have not addressed with some references to provide starting points on these more advanced topics:
\begin{itemize}
\item Covalent inhibition.
\item Free energies of mutation of protein side chains
\item nonspecific binding or multiple binding sites.
\item Complex association stoichiometry (beyond simple 1:1 association)~\cite{awesome reference}.
\item Endpoint free energy methods such as MM-PBSA~\cite{genheden2015mm} and LIE~\cite{gutierrez-de-teran2012linear}
\item Free energy methods that extract the ligand using geometric order parameters and potential of mean force methods~\cite{heinzelmann2017attachpullrelease}
\item Forcefield dependence for protein, ligand, ions, co-solvents, and co-factors. Different studies have looked at the influence of force fields and it is assumed the user has made an adequate choice for the system under study.~\cite{loeffler2018reproducibility, vassetti2019assessment, lopes2015current} 
\end{itemize}

For convenience we have also compiled a list of common acronyms, common symbols used, and some essential terminology.
\begin{tcolorbox}[title=Acronyms, colback=blue!10!white]
    \textit{CPU} - Central Processing Unit\\
    \textit{AFE} - Alchemical Free Energy\\
    \textit{BAR} - Bennett Acceptance Ratio\\
    \textit{FEP} - Free Energy Perturbation\\
    \textit{GPCR} - G-Protein Coupled Receptor\\
    \textit{GPU} - Graphics Processing Unit\\
    \textit{MBAR} - Multistate Bennett Acceptance Ratio\\
    \textit{MCSS} - Maximum Common Substructure\\
    \textit{MD} - Molecular Dynamics\\
    \textit{RBFE} - Relative Binding Free Energy\\
    \textit{RMSE} - Root Mean Square Error\\
    \textit{SAR} - Structure-Activity Relationships\\
    \textit{TI} - Thermodynamic Integration
\end{tcolorbox}
%\label{sec:abbreviations}
\begin{tcolorbox}[title=List of Symbols, colback=green!10!white]
U - Potential energy
\end{tcolorbox}

@@

%%%%%%%%%%%%
% Step 0   %
%%%%%%%%%%%%
\section{What can be expected from alchemical simulations?}
\label{sec:step0}

@@

When starting an alchemical free energy project you need to determine if an alchemical free energy calculation is the appropriate strategy by initially counting the cost of your
project: Can you even hope to tackle the problem with available resources and, if successful, will it
be worth it?
\subsection{How accurate are alchemical free energy calculations?}

Current alchemical free energy calculations seem to achieve, in favorable cases, RMS errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative. A small selection of benchmark datasets and case studies can be found in section~\ref{sec:benchmark} at the end of this document.
However, the domain of applicability is a significant concern~\cite{sherborne2016collaborating, cournia2017relative}, especially for relative calculations, which typically require a bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.

It is worth noting that the accuracy of free energy calculations is highly variable across different protein targets, and likely across different ligand chemotypes as well. For instance, FEP+ with OPLS3 achieves an RMSE of 0.62 kcal/mol for a set of 21 compounds binding to JNK1 kinase, yet only an RMSE of 1.05 kcal/mol for a set of 34 compounds binding to P38$\alpha$ kinase~\cite{harder2016opls3}. Furthermore, perturbations for the same chemotype in different pockets of the BACE enzyme gave varied errors~\cite{keranen2017acylguanidine}
\todo[inline,color={green!20}]{JDC: It's important to be clear on what error is being reported: $\Delta G$ after shifting by a constant to minimize the RMSE, unshifted $\Delta G$, $\Delta \Delta G$ of computed edges, or $\Delta \Delta G$ of all edges. Note that computed edge $\Delta \Delta G$ RMSE significantly underestimates the error of all edge RMSE: \url{https://github.com/jchodera/jacs-dataset-analysis} ASJSM: Can we incorporate your analysis in this section as an example? Maybe this is something that needs to be addressed properly in the analysis section and only referenced here?}
We thus recommend that retrospective study for a particular target and a particular chemical series be performed to establish the relevant accuracy of free energy calculations for each particular application case.
%

@@

@@

\subsection{Is my problem suitable for alchemical free energy calculations?}
Before even planning free energy calculations to study binding to a
particular target, it is important to assess what is known about the
system and its timescales and its suitability for free energy
calculations, as well as the \emph{purpose} of the calculations and
the amount of available computer resources. In some cases, predicting
accurate binding free energies for a particular target might be a
\emph{more} challenging effort than simply measuring them! This is
often the case when dealing with database screening problems, where
compounds might be easily and quickly available commercially for
testing and free energy calculations could consume a much larger
amount of resources. Free energy calculations thus typically only
appeal when (slow or costly) synthesis would be required to do
experiments or experiments are otherwise cost-prohibitive.

Sometimes free energy calculations can provide answers that are not
readily available from experiments.  For example, type II kinase
inhibitors selectively bind to different kinases in the so-called
DFG-out conformations (cite Kuriyan).  The selectivity of such
inhibitors may be attributed either to their differential binding to
different kinases in the DFG-out conformations, or to different
stability of the DFG-out conformations of different kinases.  Let
$K_C$ be the equilibrium constant between DFG-in and DFG-out
conformations of one kinase, and $K_D^\ast$ be the dissociation
constant of a type II inhibitor against this kinase, the apparent
binding constant of this inhibitor against this kinase is then
\begin{equation}
  K_D = K_D^\ast \frac{1 + K_C}{K_C}
  \label{eqn:conformational-binding}
\end{equation}

@@

\subsection{Is the expected accuracy of the computation sufficient?}
The requisite level of accuracy is another important consideration ---if the
goal is to guide lead optimization when many compounds will be
synthesized, free energy calculations can be appealing even with
accuracy in the 1--2 kcal/mol range~\cite{mobley2012perspective}, but if the number of compounds to be synthesized is very small, this accuracy may not be enough to provide
much value.

Here we include a simple estimate of the value provided by alchemical
free energy calculations in lead optimization.  Let $P(\Delta\Delta
G)$ be the probability distribution of the changes in the binding free
energies of a new set of molecules during one round of lead
optimization, and let $P(\Delta\Delta G^\dagger|\Delta\Delta G)$ be the
conditional probability of the binding free energy change computed by
the free energy calculations, $\Delta\Delta G^\dagger$, given the actual
change $\Delta\Delta G$.  The latter conditional probability can be modeled
by a normal distribution
\begin{equation}
  P(\Delta\Delta G^\dagger|\Delta\Delta G) = \frac{1}{\sqrt{2\pi\sigma^2}}
  \exp\left(-\frac{(\Delta\Delta G^\dagger - \Delta\Delta G)^2}{2\sigma^2}\right),
  \label{eqn:free-energy-distribution}
\end{equation}
where $\sigma$ signifies the accuracy of free energy calculations.
Here we assume that there is no systematic bias in the free energy
calculations, i.e., on average, the free energy change computed by
free energy calculations agrees with the actual free energy change.
%
In lead optimization guided by free energy calculations, we only
synthesize and experimentally test molecules that are predicted to
have favorable free energy changes.  We are thus interested in how
often that a molecule predicted to bind stronger actually turns out to
bind stronger.  In other words, we are interested in the conditional
probability:
\begin{equation}
  P(\Delta\Delta G<0|\Delta\Delta G^\dagger<0).
  \label{eqn:true-positive}
\end{equation}
%
For illustrative purposes, we assume that the actual changes in the
binding free energies for a set of new molecules also follow normal
distribution, that the standard deviation in the changes is $RT\ln 5$
(corresponding to a 5-fold change in the binding affinities), and that
1 in 10 new molecules have increased binding affinity ($\Delta\Delta G
\leq 0$).  Under such assumptions, the conditional probability in
Eq.~\ref{eqn:true-positive} can be easily computed.  If the accuracy
of free energy calculations is $\sigma = 1$ kcal/mol, $P(\Delta\Delta
G<0|\Delta\Delta G^\dagger<0) = 0.35$, which means that out of every
10 molecules selected for predicted favorable free energy change, on
average 3.5 molecules will have actual favorable free energy change.
In other words, selection by free energy calculations yields 3.5 times
more molecules of improved affinities than selection without free
energy calculations.
%  
Available computational resources and timescales of motion also factor
into this initial analysis. An individual free energy calculation
involves simulations at many different intermediate states (perhaps
20-40 or more) and each of these must typically be long enough to
capture the relevant motions in the system. If such motions are
microsecond events or longer, the computational cost of running 20-40
microsecond or longer simulations for each of $N$ ligands may become
prohibitive, or at least should be carefully considered. 
%
\subsection{Can I afford the calculation?}
Furthermore, are available computational resources sufficient that throughput will be
reasonable compared to needs of experimental collaborators working on
this system? How many ligands ($N$) can you afford to handle given
your computational resources? As cloud computing becomes more available, in-house GPU clusters may not be necessary if calculations are not run on a regular basis.
This analysis should be done up front, part of ``counting the cost''
of involvement in a particular project. In some cases, the analysis may conclude that free energy calculations will not be feasible for the proposed problem.
\todo[inline, color={green!20}]{ASJSM: @everyone: Might be nice to do an example calculation X perturbation on GTX 1080s on a AWS cluster cost \$\$X. Synthesis costs \$\$Y. }

Here, by ``cost'', we refer not just to financial cost of the calculations relative to experiments, but also time -- can the calculations can be run faster than experiments are done? How will the relevant resource and opportunity costs factor in? Both computation and experiment require human time, supplies (of different sorts), and equipment. In the extreme limit, for example, it would not make sense to spend a month running a binding free energy calculation if the equivalent experiment could be done in a day with resources already on hand. Such issues should be considered before deciding to conduct binding free energy calculations.
%
\subsection{Is an exploratory study what I want?}
An additional consideration is how much is known about your particular
target, ligand binding modes in the target, and any relevant motions
-- essentially, has it been studied enough to know whether it might be
suitable for free energy calculations? It is important to know if the system has hardly been studied, because should the initial calculations perform poorly, the effort may turn into an attempt to understand the relevant sampling, force field, or system preparation problems.
%
If you are unsure whether your project is feasible, one option may be
to conduct a short exploratory study to assess tractability for just a very small
number of ligands. This can be sufficient to get an initial
idea of feasibility and accuracy of the calculations for the
proposed target.

\section{How should alchemical simulations be applied to drug discovery?}
\todo[inline, color={green!20}]{ASJSM: Merge/tidy of what can I expect from my simulation, drug discovery aspects and simulation prerequisites, there are some duplications there at the moment, happy for restructuring ideas.  }

@@

\paragraph{How to deal with different binding modes?}
Generally, binding modes within congeneric series are conserved~\cite{}, however, exceptions exist~\cite{}, and weaker binding compounds are more likely to adopt different binding modes (REFS). There may also be ambiguity about the relative orientation of substituents. A known issue involves a 180 degree flip in the dihedral angle of a 6-membered aromatic ring leading to a different spatial position of ortho- or meta- substituents that otherwise should overlap within a series. The 180 degree flip of the ring may not occur enough during simulations to overcome the starting configuration bias. Another scenario may be equatorial and axial substituted saturated rings (e.g. cyclohexane derivatives). This situation may be addressed by explicitly modelling different binding modes of the same ligand and combining later computed free energy differences for different binding modes into a relative free energies of binding~\cite{kaus2015how}.

\paragraph{Have you considered sterioisomers and enationmers?}
Congeneric series can contain stereoisomers or enantiomers, making sure that you model the correct one is important. For racemates, the relative abundance of each stereoisomer is normally not known. Therefore, the experimental activity associated with just one conformation is more uncertain. However, the modeling typically uses just the bioactive conformation that best fits the active site. Clearly this introduces the chance for higher error versus experiment. Nevertheless, if all compounds in the congeneric series are racemic, originating from similar synthetic procedures with an expected similar abundance of stereoisomers, then the differences may cancel and the trend in calculated and observed binding energies may be robust. Nevertheless, we can see that care and further testing is needed in this scenario, and the quality of the predictions may suffer.

\paragraph{Conserved binding site waters can play an important role in binding free energies}
Another aspect to look out for are binding site water molecules that may form water mediated protein-ligand interactions, particularly in situations where exchange with bulk water may be slow on the simulation timescales. This happens typically in buried binding site. Overlaying multiple protein X-ray structures can identify conserved or just additional water molecules that can be useful to include. In cases where water molecules are known to play an important role in the binding, software implementations that use water sampling facilitated by Grand Canonical Monte Carlo methods may be useful. For examples see HSP90 as formed part of the D3R grand challenge 2015~\cite{mey2016blinded},A2A~\cite{brucemacdonald2018ligand}, MUP~\cite{ross2015water}, and others~\cite{michel2009energetics}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%
\begin{frontmatter}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Abstract
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Alchemical free energy calculations are a useful tool for predicting free energy differences associated with the transfer of small molecules from one environment to another.
The hallmark of these methods is the use of "bridging" potential energy functions representing \emph{alchemical} intermediate states that do not exist in as real chemical species. The data collected from these bridging alchemical thermodynamic states allows the computation of transfer free energies (or differences in transfer free energies) with orders of magnitude less simulation time than simulating the transfer process directly. 
%
While these methods are highly flexible, care must be taken in avoiding common pitfalls to ensure that computed free energy differences can be robust and reproducible for the chosen force field, and that appropriate corrections are included to permit direct comparison with experimental data.
%
In this paper, we review current best practices for several popular application domains of alchemical free energy calculations, including relative and absolute small molecule binding free energy calculations to biomolecular targets.
\todo[inline, color=green!20]{JDC: Should we migrate this document to \url{https://github.com/alchemistry}? ASJSM: Let's migrate with the submitted version, but leave it here for now?}
\end{abstract}
\end{frontmatter}

@@

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig1_what_is_alchemy/Figure.pdf}   
    \caption{\textbf{Illustration of common types of free energies differences that can be calculated using alchemical free energy methods.} A: Change in free energy due to a conformational change of the molecule. B: Partition coefficient such as $\log P$ or $\log D$ can be computed using alchemical free energy methods. Here, as an example the partition coefficient between methanol and water is shown. C: Free energy difference associated with the insertion of a molecule into a membrane. D: Effect of mutations of protein or host residues on free energies of binding.  E: Absolute free energies of binding of a small molecule to a host (e.g. protein), F: Relative free energy of binding of one molecule with respect to another, here toluene and benzyl alcohol.
    \label{fig:fig_what_is_alchemy}
    }
\end{figure}
% Prerequesites    %
\section{Prerequisites and Scope}
\label{sec:pre}
\todo[inline, color=green!20]{ASJSM: @everyone This section is up for review to also check flow with intro and rest of document.}
This guide is meant to serve as a comprehensive starting point for new practitioners for free energy calculations.  It assumes moderate experience with molecular simulation packages and concepts, however.  It is also intended to serve as a reference guide with a convenient checklist (Section~\ref{sec:checklist}) for more seasoned practitioners, to help ensure their simulation meet standard alchemical simulation and analysis practices. It can also serve as a set of best practices to ensure simulation robustness and reproducibility which reviewers may wish to consider as they evaluate papers.
This guide aims to answer the following questions:
    \item Is my problem suitable for an alchemical free energy calculation? 
    \item How do I choose the right free energy protocol and run it? 
    \item How do I analyze my simulations accurately? 
    \item What software tools are available to perform alchemical free energy calculations? 
We assume the reader possesses a basic familiarity with the principles of molecular mechanics, molecular dynamics simulations, statistical mechanics, and the biophysics of protein-ligand association.
For readers unfamiliar with these concepts that have never set up and run an equilibrium molecular dynamics simulation, we would recommend the following texts for getting started with these topics first. 
    \item Statistical mechanics and thermodynamics background~\cite{hill1956statistical, alavi2011statistical}
    \item Understanding theory of how molecular simulations work~\cite{frenkel2002understanding}
    \item Docking reviews~\cite{yuriev2013latest, meng2011molecular} and software overview~\cite{pagadala2017software} and generating valid molecular structures ~\cite{loeffler2015fesetup}
    \item Getting started with setting up and running equilibrium molecular dynamics simulations: Gromacs tutorial~\cite{lemkul2018From}, Amber tutorial~\cite{amber}, HTMD: high-throughput Molecular Dynamics for Molecular Discovery~\cite{doerr2016htmd}*, or with VMD and NAMD~\cite{ribeiro2016easy}*
References marked with an asterisk (*) indicate nonessential reading or tutorials that provide further depth and understanding.
As some of the theoretical background can seem daunting, we provide an essentials guide to the theory behind alchemical free energy calculations in Section~\ref{sec:theory}.
In the remainder of this paper, we will cover topics that are key to the preparation~(\ref{sec:step1}), choice and use of correct protocols~(\ref{sec:step2}), and finally the best practices that should be used for the analysis of AFE calculations~(\ref{sec:step3}). 
Particular focus will be given to aspects of the molecular simulations which are unique to alchemical calculations---these include the calculation of transfer free energies (hydration free energies, partition coefficients, etc.), and binding free energies (absolute and relative).
While we try to address as many methods and practices as possible, the field of free energy calculations is broad, and there are many advanced topics that are left to future best practices documents focusing on specific issues. 
Below, we provide a non-exhaustive list of topics we have not addressed with some references to provide starting points on these more advanced topics:
\item Covalent inhibition.
\item Free energies of mutation of protein side chains
\item nonspecific binding or multiple binding sites.
\item Complex association stoichiometry (beyond simple 1:1 association)~\cite{awesome reference}.
\item Endpoint free energy methods such as MM-PBSA~\cite{genheden2015mm} and LIE~\cite{gutierrez-de-teran2012linear}
\item Free energy methods that extract the ligand using geometric order parameters and potential of mean force methods~\cite{heinzelmann2017attachpullrelease}
\item Forcefield dependence for protein, ligand, ions, co-solvents, and co-factors. Different studies have looked at the influence of force fields and it is assumed the user has made an adequate choice for the system under study.~\cite{loeffler2018reproducibility, vassetti2019assessment, lopes2015current} 
For convenience we have also compiled a list of common acronyms, common symbols used, and some essential terminology.
\begin{tcolorbox}[title=Acronyms, colback=blue!10!white]
    \textit{CPU} - Central Processing Unit\\
    \textit{AFE} - Alchemical Free Energy\\
    \textit{BAR} - Bennett Acceptance Ratio\\
    \textit{FEP} - Free Energy Perturbation\\
    \textit{GPCR} - G-Protein Coupled Receptor\\
    \textit{GPU} - Graphics Processing Unit\\
    \textit{MBAR} - Multistate Bennett Acceptance Ratio\\
    \textit{MCSS} - Maximum Common Substructure\\
    \textit{MD} - Molecular Dynamics\\
    \textit{RBFE} - Relative Binding Free Energy\\
    \textit{RMSE} - Root Mean Square Error\\
    \textit{SAR} - Structure-Activity Relationships\\
    \textit{TI} - Thermodynamic Integration
\end{tcolorbox}
%\label{sec:abbreviations}
\begin{tcolorbox}[title=List of Symbols, colback=green!10!white]
U - Potential energy
\end{tcolorbox}

@@

%%%%%%%%%%%%
% Step 0   %
%%%%%%%%%%%%
\section{What can be expected from alchemical simulations?}
\label{sec:step0}

@@

When starting an alchemical free energy project you need to determine if an alchemical free energy calculation is the appropriate strategy by initially counting the cost of your
project: Can you even hope to tackle the problem with available resources and, if successful, will it
be worth it?
\subsection{How accurate are alchemical free energy calculations?}
\label{subsec:accuracy}
Current alchemical free energy calculations seem to achieve, in favorable cases, RMS errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative. A small selection of benchmark datasets and case studies can be found in section~\ref{sec:benchmark} at the end of this document.
However, the domain of applicability is a significant concern~\cite{sherborne2016collaborating, cournia2017relative}, especially for relative calculations, which typically require a bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.
It is worth noting that the accuracy of free energy calculations is highly variable across different protein targets, and likely across different ligand chemotypes as well. For instance, FEP+ with OPLS3 achieves an RMSE of 0.62 kcal/mol for a set of 21 compounds binding to JNK1 kinase, yet only an RMSE of 1.05 kcal/mol for a set of 34 compounds binding to P38$\alpha$ kinase~\cite{harder2016opls3}. Furthermore, perturbations for the same chemotype in different pockets of the BACE enzyme gave varied errors~\cite{keranen2017acylguanidine}
\todo[inline,color={green!20}]{JDC: It's important to be clear on what error is being reported: $\Delta G$ after shifting by a constant to minimize the RMSE, unshifted $\Delta G$, $\Delta \Delta G$ of computed edges, or $\Delta \Delta G$ of all edges. Note that computed edge $\Delta \Delta G$ RMSE significantly underestimates the error of all edge RMSE: \url{https://github.com/jchodera/jacs-dataset-analysis} ASJSM: Can we incorporate your analysis in this section as an example? Maybe this is something that needs to be addressed properly in the analysis section and only referenced here?}
We thus recommend that retrospective study for a particular target and a particular chemical series be performed to establish the relevant accuracy of free energy calculations for each particular application case.

@@

\label{subsec:reproducible}

@@

\subsection{Is my problem suitable for alchemical free energy calculations?}
\label{subsec:suitability}
Before even planning free energy calculations to study binding to a
particular target, it is important to assess what is known about the
system and its timescales and its suitability for free energy
calculations, as well as the \emph{purpose} of the calculations and
the amount of available computer resources. In some cases, predicting
accurate binding free energies for a particular target might be a
\emph{more} challenging effort than simply measuring them! This is
often the case when dealing with database screening problems, where
compounds might be easily and quickly available commercially for
testing and free energy calculations could consume a much larger
amount of resources. Free energy calculations thus typically only
appeal when (slow or costly) synthesis would be required to do
experiments or experiments are otherwise cost-prohibitive.
Sometimes free energy calculations can provide answers that are not
readily available from experiments.  For example, type II kinase
inhibitors selectively bind to different kinases in the so-called
DFG-out conformations (cite Kuriyan).  The selectivity of such
inhibitors may be attributed either to their differential binding to
different kinases in the DFG-out conformations, or to different
stability of the DFG-out conformations of different kinases.  Let
$K_C$ be the equilibrium constant between DFG-in and DFG-out
conformations of one kinase, and $K_D^\ast$ be the dissociation
constant of a type II inhibitor against this kinase, the apparent
binding constant of this inhibitor against this kinase is then
\begin{equation}
  K_D = K_D^\ast \frac{1 + K_C}{K_C}
  \label{eqn:conformational-binding}
\end{equation}

@@

\subsection{Is the expected accuracy of the computation sufficient?}
\label{subsec:accuracy}
The requisite level of accuracy is another important consideration ---if the
goal is to guide lead optimization when many compounds will be
synthesized, free energy calculations can be appealing even with
accuracy in the 1--2 kcal/mol range~\cite{mobley2012perspective}, but if the number of compounds to be synthesized is very small, this accuracy may not be enough to provide
much value.
Here we include a simple estimate of the value provided by alchemical
free energy calculations in lead optimization.  Let $P(\Delta\Delta
G)$ be the probability distribution of the changes in the binding free
energies of a new set of molecules during one round of lead
optimization, and let $P(\Delta\Delta G^\dagger|\Delta\Delta G)$ be the
conditional probability of the binding free energy change computed by
the free energy calculations, $\Delta\Delta G^\dagger$, given the actual
change $\Delta\Delta G$.  The latter conditional probability can be modeled
by a normal distribution
\begin{equation}
  P(\Delta\Delta G^\dagger|\Delta\Delta G) = \frac{1}{\sqrt{2\pi\sigma^2}}
  \exp\left(-\frac{(\Delta\Delta G^\dagger - \Delta\Delta G)^2}{2\sigma^2}\right),
  \label{eqn:free-energy-distribution}
\end{equation}
where $\sigma$ signifies the accuracy of free energy calculations.
Here we assume that there is no systematic bias in the free energy
calculations, i.e., on average, the free energy change computed by
free energy calculations agrees with the actual free energy change.
In lead optimization guided by free energy calculations, we only
synthesize and experimentally test molecules that are predicted to
have favorable free energy changes.  We are thus interested in how
often that a molecule predicted to bind stronger actually turns out to
bind stronger.  In other words, we are interested in the conditional
probability:
\begin{equation}
  P(\Delta\Delta G<0|\Delta\Delta G^\dagger<0).
  \label{eqn:true-positive}
\end{equation}
For illustrative purposes, we assume that the actual changes in the
binding free energies for a set of new molecules also follow normal
distribution, that the standard deviation in the changes is $RT\ln 5$
(corresponding to a 5-fold change in the binding affinities), and that
1 in 10 new molecules have increased binding affinity ($\Delta\Delta G
\leq 0$).  Under such assumptions, the conditional probability in
Eq.~\ref{eqn:true-positive} can be easily computed.  If the accuracy
of free energy calculations is $\sigma = 1$ kcal/mol, $P(\Delta\Delta
G<0|\Delta\Delta G^\dagger<0) = 0.35$, which means that out of every
10 molecules selected for predicted favorable free energy change, on
average 3.5 molecules will have actual favorable free energy change.
In other words, selection by free energy calculations yields 3.5 times
more molecules of improved affinities than selection without free
energy calculations.
Available computational resources and timescales of motion also factor
into this initial analysis. An individual free energy calculation
involves simulations at many different intermediate states (perhaps
20-40 or more) and each of these must typically be long enough to
capture the relevant motions in the system. If such motions are
microsecond events or longer, the computational cost of running 20-40
microsecond or longer simulations for each of $N$ ligands may become
prohibitive, or at least should be carefully considered. 
\subsection{Can I afford the calculation?}
\label{subsec:affordability}
Furthermore, are available computational resources sufficient that throughput will be
reasonable compared to needs of experimental collaborators working on
this system? How many ligands ($N$) can you afford to handle given
your computational resources? As cloud computing becomes more available, in-house GPU clusters may not be necessary if calculations are not run on a regular basis.
This analysis should be done up front, part of ``counting the cost''
of involvement in a particular project. In some cases, the analysis may conclude that free energy calculations will not be feasible for the proposed problem.
Here, by ``cost'', we refer not just to financial cost of the calculations relative to experiments, but also time -- can the calculations can be run faster than experiments are done? How will the relevant resource and opportunity costs factor in? Both computation and experiment require human time, supplies (of different sorts), and equipment. In the extreme limit, for example, it would not make sense to spend a month running a binding free energy calculation if the equivalent experiment could be done in a day with resources already on hand. Such issues should be considered before deciding to conduct binding free energy calculations.
\subsection{Is an exploratory study what I want?}
\label{subsec:exploration}
An additional consideration is how much is known about your particular
target, ligand binding modes in the target, and any relevant motions
-- essentially, has it been studied enough to know whether it might be
suitable for free energy calculations? It is important to know if the system has hardly been studied, because should the initial calculations perform poorly, the effort may turn into an attempt to understand the relevant sampling, force field, or system preparation problems.
If you are unsure whether your project is feasible, one option may be
to conduct a short exploratory study to assess tractability for just a very small
number of ligands. This can be sufficient to get an initial
idea of feasibility and accuracy of the calculations for the
proposed target.
\section{How should alchemical simulations be applied to drug discovery?}
\label{sec:drugdiscovery}
\todo[inline, color={green!20}]{ASJSM: Merge/tidy of what can I expect from my simulation, drug discovery aspects and simulation prerequisites, there are some duplications there at the moment, happy for restructuring ideas.  }

@@

\paragraph{How to deal with different binding modes?}
Generally, binding modes within congeneric series are conserved~\cite{}, however, exceptions exist~\cite{}, and weaker binding compounds are more likely to adopt different binding modes (REFS). There may also be ambiguity about the relative orientation of substituents. A known issue involves a 180 degree flip in the dihedral angle of a 6-membered aromatic ring leading to a different spatial position of ortho- or meta- substituents that otherwise should overlap within a series. The 180 degree flip of the ring may not occur enough during simulations to overcome the starting configuration bias. Another scenario may be equatorial and axial substituted saturated rings (e.g. cyclohexane derivatives). This situation may be addressed by explicitly modelling different binding modes of the same ligand and combining later computed free energy differences for different binding modes into a relative free energies of binding~\cite{kaus2015how}.
\paragraph{Have you considered sterioisomers and enationmers?}
Congeneric series can contain stereoisomers or enantiomers, making sure that you model the correct one is important. For racemates, the relative abundance of each stereoisomer is normally not known. Therefore, the experimental activity associated with just one conformation is more uncertain. However, the modeling typically uses just the bioactive conformation that best fits the active site. Clearly this introduces the chance for higher error versus experiment. Nevertheless, if all compounds in the congeneric series are racemic, originating from similar synthetic procedures with an expected similar abundance of stereoisomers, then the differences may cancel and the trend in calculated and observed binding energies may be robust. Nevertheless, we can see that care and further testing is needed in this scenario, and the quality of the predictions may suffer.   
\paragraph{Conserved binding site waters can play an important role in binding free energies}
Another aspect to look out for are binding site water molecules that may form water mediated protein-ligand interactions, particularly in situations where exchange with bulk water may be slow on the simulation timescales. This happens typically in buried binding site. Overlaying multiple protein X-ray structures can identify conserved or just additional water molecules that can be useful to include. In cases where water molecules are known to play an important role in the binding, software implementations that use water sampling facilitated by Grand Canonical Monte Carlo methods may be useful. For examples see HSP90 as formed part of the D3R grand challenge 2015~\cite{mey2016blinded},A2A~\cite{brucemacdonald2018ligand}, MUP~\cite{ross2015water}, and others~\cite{michel2009energetics}.
\paragraph{How to align congeneric series}
